A Graph Theoretic Approach to Problem Formulation for Multidisciplinary Design Analysis and Optimization

### 2013-09-04 reviewer_response.tex

\begin{document}
\maketitle

\section{Reviewer 1}

\rev{This idea of PSG was mentioned but was not covered in 
the paper. For completeness and value proposition sake, I would recommend to 
add at least half section to show how it will help and look like. You can 
take the simplest formulation of your choice and show how MCG and FPG will 
translate and look as a PSG. A brief discussion of advantages vs 
disadvantages will be useful. }

We agree that this was too quickly addressed in the paper. While we feel that 
discussion of the PSG in any significant depth is too big of a topic to fit into 
this paper, it does warrant a more complete description than was given in the 
original draft.

We have added sections 3.5, 4.4 in the appropriate 
parts of the paper to address the the driver nodes and driven edges which 
together distinguish a PSG from and FPG.  We have also added added section 5.5
which gives an example of a simple PSG and compares a PSG with its close relative
and XDSM diagram.

### 2013-09-04 graph_syntax.tex
adding in first PSG details

\subsection{Node and Edge Types}
  \label{ss:types}
We categorize nodes and edges 
  into distinct types in order to provide an intuitive approach to MDAO problem formulation. The three node types are
    \begin{description}
      \item[\bf{variable node:}] represents scalar or array data, inputs and outputs,
      \item[\bf{function node:}] represents the computation peformed by analysis tools,
      \item[\bf{driver node:}] represents control logic capable of managing iterations (present only in a PSG),
    \end{description}
  and the three edge types are

@@

@@

 \begin{figure}[htb!]
  \begin{center}
    \includegraphics[width=3.25in]{images/sellar_cycles}
  \end{center}
    \caption{Sellar problem graph with dashed edges indicating participation in a cycle.
\label{f:sellar cycles}
    }
\end{figure}

  Coupling cycles do not contain driver nodes in the MCG or FPG representation. 
  Solvers, optimizers, and other iterative processes are not invovled in the coupling 
  definition; however, one or more solvers will be required to build a valid PSG from a given FPG  that includes cycles. 
  Additionally, a coupling cycle has no inherent start or end point. It would be acceptable to select a
  node in the cycle as a starting point and proceed around the
  loop until the starting point is reached again. For the Sellar Problem, selecting 
  $D_1$ as the starting point would yield a problem as given in 
  Eq.~\ref{eqn:simple}.
  Larger problems can involve more complex cycles in their FPG, indicating more 
  complex coupling between analyses. For example, a cycle can involve more than 
   two analysis blocks. Multiple independent cycles could also exist, indicating 
  independent coupling relationships. Cycles can also overlap, meaning that the same analysis 
  blocks are involved in multiple different coupling cycles. All of these situations
  arise naturally as the size of problems grows, and managing this coupling may
  become difficult. In the present work, Sec.~\ref{s:example problem}
  demonstrates how building an FPG from an MCG provides an opportunity to 
  identify and potentailly reduce the number of cycles in a graph.

@@

  In a multi-fidelity problem, a given variable node may have multiple incoming 
  edges without implying a collision as defined by Eq.~\ref{e:collision}. 
  Figure \ref{f:collision_example} shows a modified version of the Sellar problem 
  with a new analysis, $D_0$, representing a low fidelity version of $D_1$. 
  In this modified version, the variable nodes with an upper indegree limit of two are highlighted in gray, and changing the upper indegree limits for the variable nodes representing $y_1$ in $D_2$, $F$, and $G_1$ avoids a collision.
In Sec.~\ref{ss:obtaining FPG} we present an algorithm for locating 
  conflicts within a given MCG such that a designer can make the necessary decisisons about each 
  one in turn. 
  \begin{figure}
    \begin{center}
      \includegraphics[width=3.25in]{images/sellar_mulfi}
    \caption{Graph of the modified Sellar problem with multi-fidelity nodes highlighed in gray.\label{f:collision_example}}
  \end{center}
  \end{figure}

  Multi-fidelity problems are always charaterized by graphs in which at least some of the variable 
  nodes that have an $\txt{deg}^-_u(v) > 1$. These problems require 
  special techniques for resolving the confliciting edges by introducing some mechanism
  to manage when each of the different fidelity analyses are 
  run \cite{march2012provably,alexandrov2001approximation,Huang_Allen_Notz_Miller_2006}.
  The specifics of this mechanism are not represented within an MCG or 
  FPG. Instead, the multi-fidelity mechanism specifics would be represented as a 
  driver in a PSG derived from a multi-fidelity FPG.

@@

>>>>

\subsection{Node and Edge Types}
  \label{ss:types}
We categorize nodes and edges 
  into distinct types in order to provide an intuitive approach to MDAO problem formulation. 
  The three node types are
    \begin{description}
      \item[\bf{variable node:}] represents scalar or array data, inputs and outputs,
      \item[\bf{function node:}] represents the computation performed by analysis tools,
      \item[\bf{driver node:}] represents control logic capable of managing iterations 
      (present only in a PSG),
    \end{description}
  and the three edge types are

@@

  \subsection{Driver Nodes and Driven Edges}
  Within this syntax, all iterative processes are represented with Driver nodes. 
  The driver node shares some basic qualities with the model node. It can have incoming 
  and outgoing function edges to and from variable nodes. For example, many optimizers take 
  convergence tolerance as input and output iteration counts and final objective value. 
  With respect to variable nodes and function edges, drivers behave exactly the same as
  model nodes and are subject to the same rules governing their use in the graph. In this
  context, driver nodes can be part of driver blocks which behave just like analysis 
  blocks.

  Driven edges are the distinguishing characteristic associated with driver nodes and
  driver blocks. Driven edges don't follow the same rules as other edges. Driven edges  
  will always connect a driver node to a variable node (in either direction), but they 
  are not considered when counting the indegree or outdegree of the variable node.
  Hence a single variable node could have multiple incoming driven edges from different 
  driver blocks. Figure \ref{f:driver block} shows a notional example where a driver block 
  represents and optimizer solving an unconstrained minimization problem.

  \begin{figure}[htb]
    \begin{center}
    \includegraphics[width=3.0in]{images/driver_block}
    \end{center}
    \vspace{-10pt}
  \caption{Example driver block with node types labeled.}
  \label{f:driver block}
  \end{figure}

@@

  \begin{figure}[htb!]
    \begin{center}
      \includegraphics[width=3.25in]{images/sellar_cycles}
    \end{center}
    \caption{Sellar problem graph with dashed edges indicating %
    participation in a cycle.\label{f:sellar cycles}}
  \end{figure}

  Coupling cycles do not contain driver nodes in the MCG or FPG representation. 
  Solvers, optimizers, and other iterative processes are not invovled in the coupling 
  definition; however, one or more solvers will be required to build a valid PSG from a 
  given FPG  that includes cycles. Additionally, a coupling cycle has no inherent 
  start or end point. It would be acceptable to select a
  node in the cycle as a starting point and proceed around the
  loop until the starting point is reached again. For the Sellar Problem, selecting 
  $D_1$ as the starting point would yield a problem as given in 
  Eq.~\ref{eqn:simple}.
  Larger problems can involve more complex cycles in their FPG, indicating more 
  complex coupling between analyses. For example, a cycle can involve more than 
   two analysis blocks. Multiple independent cycles could also exist, indicating 
  independent coupling relationships. Cycles can also overlap, meaning that the same analysis 
  blocks are involved in multiple different coupling cycles. All of these situations
  arise naturally as the size of problems grows, and managing this coupling may
  become difficult. In the present work, Sec.~\ref{s:example problem}
  demonstrates how building an FPG from an MCG provides an opportunity to 
  identify and potentially reduce the number of cycles in a graph.

@@

  In a multi-fidelity problem, a given variable node may have multiple incoming 
  edges without implying a collision as defined by Eq.~\ref{e:collision}. 
  Figure \ref{f:collision_example} shows a modified version of the Sellar problem 
  with a new analysis, $D_0$, representing a low fidelity version of $D_1$. 
  In this modified version, the variable nodes with an upper indegree limit of two 
  are highlighted in gray, and changing the upper indegree limits for the variable nodes 
  representing $y_1$ in $D_2$, $F$, and $G_1$ avoids a collision. In Sec.~\ref{ss:obtaining FPG} 
  we present an algorithm for locating conflicts within a given MCG such that a designer 
  can make the necessary decisions about each one in turn. 
  \begin{figure}
    \begin{center}
      \includegraphics[width=3.25in]{images/sellar_mulfi}
    \caption{Graph of the modified Sellar problem with multi-fidelity nodes highlighted in gray.\label{f:collision_example}}
  \end{center}
  \end{figure}

  Multi-fidelity problems are always characterized by graphs in which at least some of the variable 
  nodes that have an $\txt{deg}^-_u(v) > 1$. These problems require 
  special techniques for resolving the conflicting edges by introducing some mechanism
  to manage when each of the different fidelity analyses are 
  run \cite{march2012provably,alexandrov2001approximation,Huang_Allen_Notz_Miller_2006}.
  The specifics of this mechanism are not represented within an MCG or 
  FPG. Instead, the multi-fidelity mechanism specifics would be represented as a 
  driver in a PSG derived from a multi-fidelity FPG.

@@

\subsection{Iterative Control: Drivers}
  There are many iterative processes that are used to help solve MDAO problems such as 
  solvers, optimizers, and Design of Experiment (DOE). Abstractly, even a human picking 
  values guided by experience and trend could be viewed as an iterative process. 
  These processes represent the control structures that govern how a design problem 
  marches toward a solution but they do not fundamentally impact the definition of the 
  problem to be solved. e.g. which optimizer you choose does not change the problem you 
  are solving but will change the path you follow to solve it. Hence, as a rule, driver 
  nodes and driven edges are excluded from both the MCG and FPG which deal exclusively 
  with problem formulation. They only come into play with in the PSG where the issue of 
  how to solve a specific problem is addressed. In fact, at least one driver block is 
  required to be present in any PSG in order to select values for the design variables.

### 2013-09-04 mcg_fpg.tex
adding in first PSG details

\section{The MCG and FPG}
    \label{s:building graphs}
We now discuss how the graph-based syntax defined in Secs.~\ref{s:syntax definition} and \ref{s:graph representation}  is used to represent an MDAO problem in its entirety. This discussion is based on the MCG and the FPG.

@@

>>>>

\section{The MCG, FPG, and PSG}
    \label{s:building graphs}
We now discuss how the graph-based syntax defined in Secs.~\ref{s:syntax definition} and \ref{s:graph representation}  is used to represent an MDAO problem in its entirety. This discussion is based on the MCG and the FPG.

@@

\subsection{Problem Solution Graph}
Once a suitable FPG has been obtained the issue of how to structure a solution strategy 
can be addressed. Similar to the way an MCG can be manipulated into a FPG, an FPG can be
further manipulated into a PSG. To get from an MCG to an FPG you eliminated analysis blocks and
connection edges to reduce the problem to something solvable, but to get from an FPG to a PSG you 
add in driver blocks, driven edges to describe the structure of the solution strategy being applied 
to the problem.
